---
title: "MS_kinetics_data_analysis"
output: html_notebook
---

All input files must be in comma-separated value format (.csv). 

If you need to make a template for the ICF calculation  start at line 10. If you are performing calculations proceed to line 42

#ICF template maker
#This section is only to generate a template for the ICF file 
```{r}
library(tidyverse)

ICF_template_raw <- read_delim(file.choose(), col_names = TRUE, delim = ",", 
                             col_types="cccnnnn")
ICF_template_raw  <- rename_all(ICF_template_raw , function(x)(str_replace_all(x, " ", "_")))

ICF_template_raw  <- ICF_template_raw  %>%
  mutate(Status = str_detect(Peptide_Modified_Sequence, "[+80]")) %>%
  mutate(Status = replace(Status, Status == TRUE, "P")) %>%
  mutate(Status = replace(Status, Status == FALSE, "U"))

ICF_template_maker<- function (x){
  temp1<-x
  temp2 <- select(temp1, c("Protein_Name","Peptide_Modified_Sequence","Status"))
  temp3 <- unique(temp2)
  temp4 <- mutate( temp3, ICF=1)
  return(temp4)
}

ICF_template_maker(ICF_template_raw )->ICF_template

ICF_Template_location <- "./ICF_Template"
if(!exists(ICF_Template_location )) dir.create(ICF_Template_location )

write_delim(ICF_template, path=str_c(ICF_Template_location , "ICF_template.csv", sep="/"),
            delim = ",",
            col_names = TRUE)

```
#Kinetics calculation section 
# Input section for Kinetics calculation 
The R script is built to analyze 3 files from experiments performed under the same conditions with the same enzyme that the user desires to be analyzed and summarized together. If there are a different number of files than 3 from the same conditions with the same enzyme, the R script can be adjusted. To increase or decrease the number of replicate trials, go to lines 133 and 152. At these lines, add additional trials by adding "T4", "T5," etc. After the "T3," but before "average. Be sure to follow the formating found at those lines.  

```{r}
library(tidyverse)
## read in skyline output file

skyline_output <- read_delim(file.choose(), col_names = TRUE, delim = ",", 
                             col_types="cccnnnn")
skyline_output <- rename_all(skyline_output, function(x)(str_replace_all(x, " ", "_")))

skyline_output <- skyline_output %>%
  mutate(Status = str_detect(Peptide_Modified_Sequence, "[+80]")) %>%
  mutate(Status = replace(Status, Status == TRUE, "P")) %>%
  mutate(Status = replace(Status, Status == FALSE, "U"))


## read in correction factor output file

cf <- read_delim(file.choose(), col_names = TRUE, delim = ",", 
                 col_types="cccn")
cf <- rename_all(cf, function(x)(str_replace_all(x, " ", "_")))

min <- 10 ##Reaction time in minutes

nM <- 100 ##Reaction concentration in Nanomolar 

clean_location <- "./cleaned"
if(!exists(clean_location)) dir.create(clean_location)

output_location <- "./output"
if(!exists(output_location)) dir.create(output_location)

outfile_summary <- "Sc_Cdc14_100nM_summary.csv"

#end of input section

# Definition of Functions 


kinetics <- function(file, min, nM){
  #this function calculates kinetic values for the peptide series 
  y <- (min*60)# time in minutes
  k <- (nM*(1E-9)) #concentration in nM
  # step 1: reorganize the data 
  temp <- file
  tempsplit <- split(temp, temp$Status)
  U <- tempsplit[["U"]]
  P <- tempsplit[["P"]]
  temp <- merge(U,P, by.x=c("Protein_Name"), by.y=c("Protein_Name"), all.x=TRUE, all.y=TRUE)
  temp<-mutate(temp, "Enzyme_concentration_M" = k )# add column for enzyme concentration
  temp<-mutate(temp, "Time_sec" = y ) # add column for time
  a <- (as.numeric(temp[["Peak_Area_Corrected.x"]])) # unphosporylated peak area
  b <- (as.numeric(temp[["Peak_Area_Corrected.y"]])) # phosporylated peak area
  # step 2 calculate 
  Sr<-(b/(a+b))
  temp<-mutate(temp, "Substrate_remaining" =Sr) #calculate 
  w <-(-log((Sr))/(k*y)) # calculate kcat/km
  temp<-mutate(temp, "Kcat_Km" = w )
  temp<- select(temp, c("Protein_Name","Substrate_remaining","Kcat_Km"))
  temp<- na.omit(temp)
  # step 3 output 
  as_tibble(temp)->temp
  return(temp)
}

mycorrector <- function(x){
    # this function cleans the data and adjusts the phospho signal based on the correction factor from the lambda phosphatase data
  temp1 <- x
  temp2 <- unique(temp1)
  temp3 <- mutate(temp2, Peak_Area = Total_Area_MS1 - Total_Background_MS1)
  temp4 <- select(temp3, -c(Total_Area_MS1, Total_Background_MS1))
  temp5 <- summarize(group_by(temp4, Protein_Name, Peptide_Modified_Sequence, Status),
                     Peak_Area_Total= sum(Peak_Area))
  temp6 <- full_join(temp5, cf, by=c("Peptide_Modified_Sequence", "Status"))
  temp7 <- mutate(temp6, Peak_Area_Corrected = as.numeric(Peak_Area_Total) * as.numeric(ICF))
  temp8 <- select(temp7, -c(Peak_Area_Total, ICF))
  return(temp8)
}

Sr_summarizer <- function(x){
  # this function calculates summary statistics for the Sr values 
  temp  <- bind_rows(x, .id="file.y") 
  temp1 <- select(temp, Protein_Name, file.y, Substrate_remaining) 
  temp2 <- pivot_wider(temp1, names_from=file.y, values_from=Substrate_remaining) 
  temp3 <- arrange(temp2, Protein_Name)
  my_means <- as.matrix(select(temp3, contains("correct"))) %>% apply(1, mean)
  temp4 <- mutate(temp3, means = my_means)
  my_sd <- as.matrix(select(temp3, contains("correct"))) %>% apply(1, sd)
  temp5 <- mutate(temp4, sd = my_sd)
  name_row2 <- c("Peptide_name","T1_Sr","T2_Sr","T3_Sr","average_Sr","std_Sr")
  colnames(temp5)<- name_row2
  temp5[["average_Sr"]]->average
  temp5[["std_Sr"]]-> std
  cv <- std/average
  temp6<- mutate(temp5, "CV_Sr"=cv)
  return(temp6)
}

kinetics_summarizer <- function(x){
  # this function calculates summary statistics for the Sr values 
  temp  <- bind_rows(x, .id="file.y") 
  temp1 <- select(temp, Protein_Name, file.y, Kcat_Km) 
  temp2 <- pivot_wider(temp1, names_from=file.y, values_from=Kcat_Km) 
  temp3 <- arrange(temp2, Protein_Name)
  my_means <- as.matrix(select(temp3, contains("correct"))) %>% apply(1, mean)
  temp4 <- mutate(temp3, means = my_means)
  my_sd <- as.matrix(select(temp3, contains("correct"))) %>% apply(1, sd)
  temp5 <- mutate(temp4, sd = my_sd)
  name_row2 <- c("Peptide_name","T1_kcat_km","T2_kcat_km","T3_kcat_km","average_kcat_KM","std_kcat_KM")
  colnames(temp5)<- name_row2
  temp5[["average_kcat_KM"]]->average
  temp5[["std_kcat_KM"]]-> std
  cv <- std/average
  temp6<- mutate(temp5, "CV_kcat_KM"=cv)
  return(temp6)
}


# Data cleaning and analysis


skyline_output <- select(skyline_output,-Retention_Time) # remove columns 
skyline_split <- split(skyline_output, skyline_output$Replicate) # split the skyline file into individual replicates 

all_correct <- lapply(skyline_split, mycorrector) # apply the my corrector function across all replicates

for(i in 1:length(all_correct)){
  write_delim(all_correct[[i]], path=str_c(clean_location, "/correct_", names(all_correct)[i], "_clean.csv", sep=""), delim=",") # output cleaned files 
}

 file_list1 <- list.files(path=clean_location, pattern="*clean.csv", full.names=TRUE) # create list of all .csv files in folder
  # read in each .csv file in file_list and create a data frame with the same name as the .csv file
 
df_list <- lapply(file_list1, read_delim, delim=",", col_names=TRUE) # add the file name into each individual file as a column in list 1
names(df_list) <- file_list1

calculated<- lapply(df_list, kinetics, min, nM) # apply kinetics function each individual file in the sequence 

Kcat.Km <- kinetics_summarizer(calculated) # calculate summary statistics for kcat/km

Sr <- Sr_summarizer(calculated)# calculate summary statistics for Sr

summary<- left_join(Kcat.Km,Sr,by="Peptide_name") # combine into a single tibble


write_delim(summary, path=str_c(output_location, outfile_summary, sep="/"),
            delim = ",",
            col_names = TRUE)

```
